---
title: "R_Homework"
author: Angela Bunning
date: 2.23.17
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load in data to my project: 
`fang_et_al_genotypes.txt`:
```{r}
genotype_data_raw <- read.table(file = "fang_et_al_genotypes.txt", as.is = TRUE)
```

`snp_position.txt`:
```{r}
snp_position_data_raw <- read.delim(file = "snp_position.txt", as.is = TRUE)
```

#Data Inspection:

`fang_et_al_genotypes.txt`:  

File size: 10.5 MB

Structure:
```{r}
str(genotype_data_raw)
```

* It is a data frame with 2783 observations and 986 variables total


Number of columns and rows:
```{r}
ncol(genotype_data_raw)      #986 columns
nrow(genotype_data_raw)      #2783 rows
```

Names of each variable:
```{r}
names_genotype <- names(genotype_data_raw)
names_genotype
```

Summary:
```{r}
genotype_summary <- summary(genotype_data_raw)
genotype_summary
```

`snp_position.txt`

File size: 80.8 KB

Structure: 
```{r}
str(snp_position_data_raw)
```

* data frame with 983 observations of 15 variables 

Number of rows and columns: 
```{r}
rows_snp_data <- nrow(snp_position_data_raw) # 983 rows
columns_snp_data <- ncol(snp_position_data_raw) # 15 columns 
```

Names of each variable:
```{r}
names_snps <- names(snp_position_data_raw)
names_snps
```

Summary:
```{r}
snp_summary <- summary(snp_position_data_raw)
snp_summary
```

# Data Proccessing 

First, separate out the groups of interest we want for maize (ZMMIL, ZMMLR, and ZMMMR) and teosinte (ZMPBA, ZMPIL, and ZMPJA) from `fang_et_al_genotypes.txt`:

```{r}
library(dplyr)
require(dplyr)
```

For maize:
```{r}
genotypes_by_groups_maize <- filter(genotype_data_raw, V3 == "ZMMIL" | V3 == "ZMMLR" | V3 == "ZMMMR")
genotypes_by_groups_maize
```

For teosinte: 
```{r}
genotypes_by_groups_teosinte <- filter(genotype_data_raw, V3 == "ZMPBA" | V3 == "ZMPIL" | V3 == "ZMPJA")
genotypes_by_groups_teosinte
```




