---
title: "R_Homework"
author: Angela Bunning
date: 2.23.17
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load in data to my project: 
`fang_et_al_genotypes.txt`:
```{r}
genotype_data_raw <- read.table(file = "fang_et_al_genotypes.txt", header = TRUE)

```

`snp_position.txt`:
```{r}
snp_position_data_raw <- read.delim(file = "snp_position.txt")
```


#Data Inspection:

`fang_et_al_genotypes.txt`:  

File size: 10.5 MB

Structure:
```{r}
str(genotype_data_raw)
```

* It is a data frame with 2783 observations and 986 variables total


Number of columns and rows:
```{r}
ncol(genotype_data_raw)      #986 columns
nrow(genotype_data_raw)      #2783 rows
```

Names of each variable:
```{r}
names_genotype <- names(genotype_data_raw)
names_genotype
```

Summary:
```{r}
genotype_summary <- summary(genotype_data_raw)
genotype_summary
```

`snp_position.txt`

File size: 80.8 KB

Structure: 
```{r}
str(snp_position_data_raw)
```

* data frame with 983 observations of 15 variables 

Number of rows and columns: 
```{r}
rows_snp_data <- nrow(snp_position_data_raw) # 983 rows
columns_snp_data <- ncol(snp_position_data_raw) # 15 columns 
```

Names of each variable:
```{r}
names_snps <- names(snp_position_data_raw)
names_snps
```

Summary:
```{r}
snp_summary <- summary(snp_position_data_raw)
snp_summary
```

# Data Proccessing 

* First, separate out the groups of interest we want for maize (ZMMIL, ZMMLR, and ZMMMR) and teosinte (ZMPBA, ZMPIL, and ZMPJA) from `fang_et_al_genotypes.txt`:

```{r, message=FALSE}
library(dplyr)
```

**maize:**
```{r}
genotypes_by_groups_maize <- filter(genotype_data_raw, Group == "ZMMIL" | Group == "ZMMLR" | Group == "ZMMMR")

head(genotypes_by_groups_maize)
```

**teosinte:**
```{r}
genotypes_by_groups_teosinte <- filter(genotype_data_raw, Group == "ZMPBA" | Group == "ZMPIL" | Group == "ZMPJA")

head(genotypes_by_groups_teosinte)
```


* Transpose data for both maize and teosinte:

**maize:**
```{r}
genotypes_maize_transposed <- t(genotypes_by_groups_maize)

head(genotypes_maize_transposed)
```

**teosinte:**
```{r}
genotypes_teosinte_transposed <- t(genotypes_by_groups_teosinte)

head(genotypes_teosinte_transposed)
```

* Only need certain columns from the `snp_position.txt` file for the analysis so we can cut those out by using the `select` function in dplyr:
```{r}
snp_columns_needed <- select(snp_position_data_raw, SNP_ID, Chromosome, Position)

head(snp_columns_needed)
```


* Rename row names in `snp_position.txt`:
```{r}
rownames(snp_columns_needed) <- snp_columns_needed$SNP_ID
```

* Now we need to join the `snp_position.txt` with each of the maize group transposed and teosinte group transposed data files:

**maize:**
```{r}
joined_maize_genotypes_snp <- merge(snp_columns_needed, genotypes_maize_transposed, by ="row.names", all = FALSE)

joined_maize_genotypes_snp <- subset(joined_maize_genotypes_snp, select = -c(2,5:15))
```

**teosinte:**
```{r}
joined_teosinte_genotypes_snp <- merge(snp_columns_needed, genotypes_teosinte_transposed, by = "row.names", all = FALSE)

joined_teosinte_genotypes_snp <- subset(joined_teosinte_genotypes_snp, select = -c(2,5:15))

```

* For all **maize** files that are chromosome specific (chromosome 1-10), with **acending** chromosome position values, and have missing data encoded by a **?** : 

**maize** chromomsome 1:
```{r}
chromosome1_maize_genotypes_snp <- filter(joined_maize_genotypes_snp, Chromosome == 1)

chromosome1_maize_genotypes_snp_arrange_ascending <- arrange(chromosome1_maize_genotypes_snp, Position)

write.csv(chromosome1_maize_genotypes_snp_arrange_ascending, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/maize_chr_ascend_question/chr1_maize_ascending_question.csv", row.names=F)
```

**maize** chromosome 2:
```{r}
chromosome2_maize_genotypes_snp <- filter(joined_maize_genotypes_snp, Chromosome == 2)

chromosome2_maize_genotypes_snp_arrange_ascending <- arrange(chromosome2_maize_genotypes_snp, Position)

write.csv(chromosome2_maize_genotypes_snp_arrange_ascending, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/maize_chr_ascend_question/chr2_maize_ascending_question.csv", row.names=F)
```

**maize** chromosome 3: 
```{r}
chromosome3_maize_genotypes_snp <- filter(joined_maize_genotypes_snp, Chromosome == 3)

chromosome3_maize_genotypes_snp_arrange_ascending <- arrange(chromosome3_maize_genotypes_snp, Position)

write.csv(chromosome3_maize_genotypes_snp_arrange_ascending, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/maize_chr_ascend_question/chr3_maize_ascending_question.csv", row.names=F)
```

**maize** chromosome 4:
```{r}
chromosome4_maize_genotypes_snp <- filter(joined_maize_genotypes_snp, Chromosome == 4)

chromosome4_maize_genotypes_snp_arrange_ascending <- arrange(chromosome2_maize_genotypes_snp, Position)

write.csv(chromosome4_maize_genotypes_snp_arrange_ascending, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/maize_chr_ascend_question/chr4_maize_ascending_question.csv", row.names=F)
```

**maize** chromosome 5:
```{r}
chromosome5_maize_genotypes_snp <- filter(joined_maize_genotypes_snp, Chromosome == 5)

chromosome5_maize_genotypes_snp_arrange_ascending <- arrange(chromosome5_maize_genotypes_snp, Position)

write.csv(chromosome5_maize_genotypes_snp_arrange_ascending, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/maize_chr_ascend_question/chr5_maize_ascending_question.csv", row.names=F)
```


**maize** chromosome 6:
```{r}
chromosome6_maize_genotypes_snp <- filter(joined_maize_genotypes_snp, Chromosome == 6)

chromosome6_maize_genotypes_snp_arrange_ascending <- arrange(chromosome6_maize_genotypes_snp, Position)

write.csv(chromosome6_maize_genotypes_snp_arrange_ascending, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/maize_chr_ascend_question/chr6_maize_ascending_question.csv", row.names=F)
```

**maize** chromosome 7:
```{r}
chromosome7_maize_genotypes_snp <- filter(joined_maize_genotypes_snp, Chromosome == 7)

chromosome7_maize_genotypes_snp_arrange_ascending <- arrange(chromosome7_maize_genotypes_snp, Position)

write.csv(chromosome7_maize_genotypes_snp_arrange_ascending, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/maize_chr_ascend_question/chr7_maize_ascending_question.csv", row.names=F)
```

**maize** chromosome 8:
```{r}
chromosome8_maize_genotypes_snp <- filter(joined_maize_genotypes_snp, Chromosome == 8)

chromosome8_maize_genotypes_snp_arrange_ascending <- arrange(chromosome8_maize_genotypes_snp, Position)

write.csv(chromosome8_maize_genotypes_snp_arrange_ascending, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/maize_chr_ascend_question/chr8_maize_ascending_question.csv", row.names=F)
```


**maize** chromosome 9:
```{r}
chromosome9_maize_genotypes_snp <- filter(joined_maize_genotypes_snp, Chromosome == 9)

chromosome9_maize_genotypes_snp_arrange_ascending <- arrange(chromosome9_maize_genotypes_snp, Position)

write.csv(chromosome9_maize_genotypes_snp_arrange_ascending, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/maize_chr_ascend_question/chr9_maize_ascending_question.csv", row.names=F)
```


**maize** chromosome 10:
```{r}
chromosome10_maize_genotypes_snp <- filter(joined_maize_genotypes_snp, Chromosome == 10)

chromosome10_maize_genotypes_snp_arrange_ascending <- arrange(chromosome10_maize_genotypes_snp, Position)

write.csv(chromosome10_maize_genotypes_snp_arrange_ascending, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/maize_chr_ascend_question/chr10_maize_ascending_question.csv", row.names=F)
```


* For all **teosinte** files that are chromosome specific (chromosome 1-10), with **acending** chromosome position values, and have missing data encoded by a **?** : 

**teosinte** chromosome 1: 
```{r}
chromosome1_teosinte_genotypes_snp <- filter(joined_teosinte_genotypes_snp, Chromosome == 1)

chromosome1_teosinte_genotypes_snp_arrange_ascending <- arrange(chromosome1_teosinte_genotypes_snp, Position)

write.csv(chromosome1_teosinte_genotypes_snp_arrange_ascending, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/teosinte_chr_ascend_question/chr1_teosinte_ascending_question.csv", row.names=F)
```

**teosinte** chromosome 2: 
```{r}
chromosome2_teosinte_genotypes_snp<- filter(joined_teosinte_genotypes_snp, Chromosome == 2)

chromosome2_teosinte_genotypes_snp_arrange_ascending <- arrange(chromosome2_teosinte_genotypes_snp, Position)

write.csv(chromosome2_teosinte_genotypes_snp_arrange_ascending, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/teosinte_chr_ascend_question/chr2_teosinte_ascending_question.csv", row.names=F)
```


**teosinte** chromosome 3:
```{r}
chromosome3_teosinte_genotypes_snp<- filter(joined_teosinte_genotypes_snp, Chromosome == 3)

chromosome3_teosinte_genotypes_snp_arrange_ascending <- arrange(chromosome3_teosinte_genotypes_snp, Position)

write.csv(chromosome3_teosinte_genotypes_snp_arrange_ascending, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/teosinte_chr_ascend_question/chr3_teosinte_ascending_question.csv", row.names=F)
```


**teosinte** chromosome 4:
```{r}
chromosome4_teosinte_genotypes_snp<- filter(joined_teosinte_genotypes_snp, Chromosome == 4)

chromosome4_teosinte_genotypes_snp_arrange_ascending <- arrange(chromosome4_teosinte_genotypes_snp, Position)

write.csv(chromosome4_teosinte_genotypes_snp_arrange_ascending, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/teosinte_chr_ascend_question/chr4_teosinte_ascending_question.csv", row.names=F)
```

**teosinte** chromosome 5:
```{r}
chromosome5_teosinte_genotypes_snp<- filter(joined_teosinte_genotypes_snp, Chromosome == 5)

chromosome5_teosinte_genotypes_snp_arrange_ascending <- arrange(chromosome5_teosinte_genotypes_snp, Position)

write.csv(chromosome5_teosinte_genotypes_snp_arrange_ascending, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/teosinte_chr_ascend_question/chr5_teosinte_ascending_question.csv", row.names=F)
```

**teosinte** chromosome 6:
```{r}
chromosome6_teosinte_genotypes_snp<- filter(joined_teosinte_genotypes_snp, Chromosome == 6)

chromosome6_teosinte_genotypes_snp_arrange_ascending <- arrange(chromosome6_teosinte_genotypes_snp, Position)

write.csv(chromosome6_teosinte_genotypes_snp_arrange_ascending, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/teosinte_chr_ascend_question/chr6_teosinte_ascending_question.csv", row.names=F)
```

**teosinte** chromosome 7:
```{r}
chromosome7_teosinte_genotypes_snp<- filter(joined_teosinte_genotypes_snp, Chromosome == 7)

chromosome7_teosinte_genotypes_snp_arrange_ascending <- arrange(chromosome7_teosinte_genotypes_snp, Position)

write.csv(chromosome7_teosinte_genotypes_snp_arrange_ascending, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/teosinte_chr_ascend_question/chr7_teosinte_ascending_question.csv", row.names=F)
```

**teosinte** chromosome 8:
```{r}
chromosome8_teosinte_genotypes_snp<- filter(joined_teosinte_genotypes_snp, Chromosome == 8)

chromosome8_teosinte_genotypes_snp_arrange_ascending <- arrange(chromosome8_teosinte_genotypes_snp, Position)

write.csv(chromosome8_teosinte_genotypes_snp_arrange_ascending, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/teosinte_chr_ascend_question/chr8_teosinte_ascending_question.csv", row.names=F)
```

**teosinte** chromosome 9:
```{r}
chromosome9_teosinte_genotypes_snp<- filter(joined_teosinte_genotypes_snp, Chromosome == 9)

chromosome9_teosinte_genotypes_snp_arrange_ascending <- arrange(chromosome9_teosinte_genotypes_snp, Position)

write.csv(chromosome9_teosinte_genotypes_snp_arrange_ascending, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/teosinte_chr_ascend_question/chr9_teosinte_ascending_question.csv", row.names=F)
```

**teosinte** chromosome 10:
```{r}
chromosome10_teosinte_genotypes_snp<- filter(joined_teosinte_genotypes_snp, Chromosome == 10)

chromosome10_teosinte_genotypes_snp_arrange_ascending <- arrange(chromosome10_teosinte_genotypes_snp, Position)

write.csv(chromosome10_teosinte_genotypes_snp_arrange_ascending, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/teosinte_chr_ascend_question/chr10_teosinte_ascending_question.csv", row.names=F)
```


For all the **maize** files that are chromosome specific (chromosome 1-10), with **decending** chromosome position values, and missing data encoded by a **-** :

**maize** chromosome 1:
```{r}
chromosome1_maize_genotypes_snp_arrange_descending <- arrange(chromosome1_maize_genotypes_snp, desc(Position))

chromosome1_maize_decending_dash <- as.data.frame(lapply(chromosome1_maize_genotypes_snp_arrange_descending, FUN = function(x)as.character(gsub("\\?", "-", x))))

write.csv(chromosome1_maize_decending_dash, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/maize_chr_decend_dash/chr1_maize_decending_dash.csv", row.names=F)
```

**maize** chromosome 2:
```{r}
chromosome2_maize_genotypes_snp_arrange_descending <- arrange(chromosome2_maize_genotypes_snp, desc(Position))

chromosome2_maize_decending_dash <- as.data.frame(lapply(chromosome2_maize_genotypes_snp_arrange_descending, FUN = function(x)as.character(gsub("\\?", "-", x))))

write.csv(chromosome2_maize_decending_dash, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/maize_chr_decend_dash/chr2_maize_decending_dash.csv", row.names=F)
```

**maize** chromosome 3:
```{r}
chromosome3_maize_genotypes_snp_arrange_descending <- arrange(chromosome3_maize_genotypes_snp, desc(Position))

chromosome3_maize_decending_dash <- as.data.frame(lapply(chromosome3_maize_genotypes_snp_arrange_descending, FUN = function(x)as.character(gsub("\\?", "-", x))))

write.csv(chromosome3_maize_decending_dash, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/maize_chr_decend_dash/chr3_maize_decending_dash.csv", row.names=F)
```



**maize** chromosome 4:
```{r}
chromosome4_maize_genotypes_snp_arrange_descending <- arrange(chromosome4_maize_genotypes_snp, desc(Position))

chromosome4_maize_decending_dash <- as.data.frame(lapply(chromosome4_maize_genotypes_snp_arrange_descending, FUN = function(x)as.character(gsub("\\?", "-", x))))

write.csv(chromosome4_maize_decending_dash, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/maize_chr_decend_dash/chr4_maize_decending_dash.csv", row.names=F)
```


**maize** chromosome 5:
```{r}
chromosome5_maize_genotypes_snp_arrange_descending <- arrange(chromosome5_maize_genotypes_snp, desc(Position))

chromosome5_maize_decending_dash <- as.data.frame(lapply(chromosome5_maize_genotypes_snp_arrange_descending, FUN = function(x)as.character(gsub("\\?", "-", x))))

write.csv(chromosome5_maize_decending_dash, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/maize_chr_decend_dash/chr5_maize_decending_dash.csv", row.names=F)
```


**maize** chromosome 6:
```{r}
chromosome6_maize_genotypes_snp_arrange_descending <- arrange(chromosome6_maize_genotypes_snp, desc(Position))

chromosome6_maize_decending_dash <- as.data.frame(lapply(chromosome6_maize_genotypes_snp_arrange_descending, FUN = function(x)as.character(gsub("\\?", "-", x))))

write.csv(chromosome6_maize_decending_dash, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/maize_chr_decend_dash/chr6_maize_decending_dash.csv", row.names=F)
```

**maize** chromosome 7:
```{r}
chromosome7_maize_genotypes_snp_arrange_descending <- arrange(chromosome7_maize_genotypes_snp, desc(Position))

chromosome7_maize_decending_dash <- as.data.frame(lapply(chromosome7_maize_genotypes_snp_arrange_descending, FUN = function(x)as.character(gsub("\\?", "-", x))))

write.csv(chromosome7_maize_decending_dash, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/maize_chr_decend_dash/chr7_maize_decending_dash.csv", row.names=F)
```

**maize** chromosome 8: 
```{r}
chromosome8_maize_genotypes_snp_arrange_descending <- arrange(chromosome8_maize_genotypes_snp, desc(Position))

chromosome8_maize_decending_dash <- as.data.frame(lapply(chromosome8_maize_genotypes_snp_arrange_descending, FUN = function(x)as.character(gsub("\\?", "-", x))))

write.csv(chromosome8_maize_decending_dash, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/maize_chr_decend_dash/chr8_maize_decending_dash.csv", row.names=F)
```

**maize** chromosome 9:
```{r}
chromosome9_maize_genotypes_snp_arrange_descending <- arrange(chromosome9_maize_genotypes_snp, desc(Position))

chromosome9_maize_decending_dash <- as.data.frame(lapply(chromosome9_maize_genotypes_snp_arrange_descending, FUN = function(x)as.character(gsub("\\?", "-", x))))

write.csv(chromosome9_maize_decending_dash, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/maize_chr_decend_dash/chr9_maize_decending_dash.csv", row.names=F)
```


**maize** chromosome 10:
```{r}
chromosome10_maize_genotypes_snp_arrange_descending <- arrange(chromosome10_maize_genotypes_snp, desc(Position))

chromosome10_maize_decending_dash <- as.data.frame(lapply(chromosome10_maize_genotypes_snp_arrange_descending, FUN = function(x)as.character(gsub("\\?", "-", x))))

write.csv(chromosome10_maize_decending_dash, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/maize_chr_decend_dash/chr10_maize_decending_dash.csv", row.names=F)
```

For all the **teosinte** files that are chromosome specific (chromosome 1-10), with **decending** chromosome position values, and missing data encoded by a **-** :

**teosinte** chromosome 1:
```{r}
chromosome1_teosinte_genotypes_snp_arrange_descending <- arrange(chromosome1_teosinte_genotypes_snp, desc(Position))

chromosome1_teosinte_decending_dash <- as.data.frame(lapply(chromosome1_teosinte_genotypes_snp_arrange_descending, FUN = function(x)as.character(gsub("\\?", "-", x))))

write.csv(chromosome1_teosinte_decending_dash, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/teosinte_chr_decend_dash/chr1_teosinte_decending_dash.csv", row.names=F)
```

**teostine** chromosome 2:
```{r}
chromosome2_teosinte_genotypes_snp_arrange_descending <- arrange(chromosome2_teosinte_genotypes_snp, desc(Position))

chromosome2_teosinte_decending_dash <- as.data.frame(lapply(chromosome2_teosinte_genotypes_snp_arrange_descending, FUN = function(x)as.character(gsub("\\?", "-", x))))

write.csv(chromosome2_teosinte_decending_dash, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/teosinte_chr_decend_dash/chr2_teosinte_decending_dash.csv", row.names=F)
```

**teosinte** chromosome 3:
```{r}
chromosome3_teosinte_genotypes_snp_arrange_descending <- arrange(chromosome3_teosinte_genotypes_snp, desc(Position))

chromosome3_teosinte_decending_dash <- as.data.frame(lapply(chromosome3_teosinte_genotypes_snp_arrange_descending, FUN = function(x)as.character(gsub("\\?", "-", x))))

write.csv(chromosome3_teosinte_decending_dash, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/teosinte_chr_decend_dash/chr3_teosinte_decending_dash.csv", row.names=F)
```

**teosinte** chromosome 4:
```{r}
chromosome4_teosinte_genotypes_snp_arrange_descending <- arrange(chromosome4_teosinte_genotypes_snp, desc(Position))

chromosome4_teosinte_decending_dash <- as.data.frame(lapply(chromosome4_teosinte_genotypes_snp_arrange_descending, FUN = function(x)as.character(gsub("\\?", "-", x))))

write.csv(chromosome4_teosinte_decending_dash, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/teosinte_chr_decend_dash/chr4_teosinte_decending_dash.csv", row.names=F)
```

**teosinte** chromosome 5:
```{r}
chromosome5_teosinte_genotypes_snp_arrange_descending <- arrange(chromosome5_teosinte_genotypes_snp, desc(Position))

chromosome5_teosinte_decending_dash <- as.data.frame(lapply(chromosome5_teosinte_genotypes_snp_arrange_descending, FUN = function(x)as.character(gsub("\\?", "-", x))))

write.csv(chromosome5_teosinte_decending_dash, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/teosinte_chr_decend_dash/chr5_teosinte_decending_dash.csv", row.names=F)
```

**teosinte** chromosome 6:
```{r}
chromosome6_teosinte_genotypes_snp_arrange_descending <- arrange(chromosome6_teosinte_genotypes_snp, desc(Position))

chromosome6_teosinte_decending_dash <- as.data.frame(lapply(chromosome6_teosinte_genotypes_snp_arrange_descending, FUN = function(x)as.character(gsub("\\?", "-", x))))

write.csv(chromosome6_teosinte_decending_dash, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/teosinte_chr_decend_dash/chr6_teosinte_decending_dash.csv", row.names=F)
```

**teosinte** chromosome 7:
```{r}
chromosome7_teosinte_genotypes_snp_arrange_descending <- arrange(chromosome7_teosinte_genotypes_snp, desc(Position))

chromosome7_teosinte_decending_dash <- as.data.frame(lapply(chromosome7_teosinte_genotypes_snp_arrange_descending, FUN = function(x)as.character(gsub("\\?", "-", x))))

write.csv(chromosome7_teosinte_decending_dash, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/teosinte_chr_decend_dash/chr7_teosinte_decending_dash.csv", row.names=F)
```

**teosinte** chromosome 8:
```{r}
chromosome8_teosinte_genotypes_snp_arrange_descending <- arrange(chromosome8_teosinte_genotypes_snp, desc(Position))

chromosome8_teosinte_decending_dash <- as.data.frame(lapply(chromosome8_teosinte_genotypes_snp_arrange_descending, FUN = function(x)as.character(gsub("\\?", "-", x))))

write.csv(chromosome8_teosinte_decending_dash, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/teosinte_chr_decend_dash/chr8_teosinte_decending_dash.csv", row.names=F)
```

**teosinte** chromosome 9:
```{r}
chromosome9_teosinte_genotypes_snp_arrange_descending <- arrange(chromosome9_teosinte_genotypes_snp, desc(Position))

chromosome9_teosinte_decending_dash <- as.data.frame(lapply(chromosome9_teosinte_genotypes_snp_arrange_descending, FUN = function(x)as.character(gsub("\\?", "-", x))))

write.csv(chromosome9_teosinte_decending_dash, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/teosinte_chr_decend_dash/chr9_teosinte_decending_dash.csv", row.names=F)
```


**teosinte** chromosome 10:
```{r}
chromosome10_teosinte_genotypes_snp_arrange_descending <- arrange(chromosome10_teosinte_genotypes_snp, desc(Position))

chromosome10_teosinte_decending_dash <- as.data.frame(lapply(chromosome10_teosinte_genotypes_snp_arrange_descending, FUN = function(x)as.character(gsub("\\?", "-", x))))

write.csv(chromosome10_teosinte_decending_dash, "~/Desktop/BCB546_Assignments/R_Assignment/Final_files/teosinte_chr_decend_dash/chr10_teosinte_decending_dash.csv", row.names=F)
```

# Part II

* Reshape raw data with `melt` using `reshape2`:
```{r}
library(reshape2)
tidy_fang <- melt(genotype_data_raw, c(1,3), c(4:986))
head(tidy_fang)
tidy_data_teosinte <- melt(joined_teosinte_genotypes_snp)
tidy_snp <- melt(snp_position_data_raw, "Chromosome", "SNP_ID")
```

* Plot total SNPs per chromosome: 
```{r}
library(ggplot2)
total_snps_per_chromosome <- ggplot(data = tidy_snp, aes(x = Chromosome)) + geom_bar() + ylab("SNPs per chromosome") + xlab("Chromosome Number")

total_snps_per_chromosome
```

* It seems that the most SNPs map to chromosome 1 between both maize and teosinte. 


* Plot Missing data and amount of heterozygosity
```{r}
homo_het <- function(x){
  if(x == "?/?")
    return("NA")
  else if (x == "G/G" | x == "A/A" | x == "C/C" | x == "T/T") 
    return("homo")
  else
    return("het")
}
tidy_fang$homo_het <- as.character(lapply(tidy_fang$value, FUN = homo_het))
tidy_fang
```

```{r}
ZMMIL <- tidy_fang$homo_het[tidy_fang$Group == "ZMMIL"]

ZMMLR <- tidy_fang$homo_het[tidy_fang$Group == "ZMMLR"]

ZMMMR <- tidy_fang$homo_het[tidy_fang$Group == "ZMMMR"]

ZMPBA <- tidy_fang$homo_het[tidy_fang$Group == "ZMPBA"]

ZMPIL <- tidy_fang$homo_het[tidy_fang$Group == "ZMPIL"]

ZMPJA <- tidy_fang$homo_het[tidy_fang$Group == "ZMPJA"]


ZMMIL <- c(sum(c(ZMMIL == "homo")),sum(c(ZMMIL == "het")),sum(c(ZMMIL == "NA")))

ZMMLR <-c(sum(c(ZMMLR == "homo")), sum(c(ZMMLR == "het")), sum(c(ZMMLR == "NA")))

ZMMMR <- c(sum(c(ZMMMR == "homo")), sum(c(ZMMMR == "het")), sum(c(ZMMMR == "NA")))

ZMPBA <- c(sum(c(ZMPBA == "homo")), sum(c(ZMPBA == "het")), sum(c(ZMPBA == "NA")))

ZMPIL <- c(sum(c(ZMPIL == "homo")),sum(c(ZMPIL == "het")),sum(c(ZMPIL == "NA")))

ZMPJA <- c(sum(c(ZMPJA == "homo")), sum(c(ZMPJA == "het")), sum(c(ZMPJA == "NA")))

het_data.frame <- data.frame(cbind(ZMMIL, ZMMLR, ZMMMR, ZMPBA, ZMPIL, ZMPJA))
grps <- c("ZMMIL", "ZMMLR", "ZMMMR", "ZMPBA","ZMPIL", "ZMPJA")
het_df.melt <- melt(het_data.frame)
het_df.melt$type <- c("homo", "het", "NA")
het_df.melt
ggplot(data = het_df.melt, aes(x = variable, y = value, fill = type)) + geom_bar(stat = "identity", position= "dodge") + xlab("Group") + ylab("Number of SNPs")

```



